---
title: "README"
author: "Micah E. Hirsch, Ph.D. mehirsch@bu.edu"
date: "2025-06-25"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# aRtic

This package is designed to preprocess articulatory kinematic data collected from EMA systems. It is currently in beta, designed to support primary use cases with limited functionality.

Further enhancements and expanded features are planned based on user feedback. Ideal for users comfortable with scripting in R. A shiny app may be developed in the future

## Installation

You can install the latest development version of the package from Github.

```{r, eval = F}

install.packages("remotes") # if you don't have remotes installed
remotes::install_github("SteppLab/aRtic")

```

## Usage

Here is a basic example of how to use the package. For more information on function syntax, please check the documentation under the /man folder.

```{r, eval = F}

library(aRtic)

# Load Your Data

bite_plane_data <- load_tsv("path_to_your_data.tsv") # Bite Plane Calibration Recording
sensor_data <- load_tsv("path_to_your_data.tsv") # Sensor Recording 
palate_trace_data <- load_tsv("path_to_your_data.tsv") # Palate Trace Recoridng (optional)

# Define referent, biteplane, and palate sensor indices
## For the current version, three referent sensors are required (2 on the left and right mastoids and 1 on the incisor)

ref_idx <- c(1, 2, 3)
bp_idx <- c(5, 6, 7)
pl_idx <- 8

# Define the coordinate plane from the Bite Plane Recording
coord <- define_coord(bite_plane_data[[1]], ref_idx, bp_idx)

# Interpolate over missing data and filter sensor recording
filtered <- interp_filter(sensor_data[[1]], ref_idx)

# Rotate sensor data to be aligned with the coordinate plane and correct for head movement
corrected <- correct_mov(filtered, coord[[1]], ref_idx, coord[[2]], coord[[3]])

# Estimate X, Y, and Z coordinates of palate trace (optional)
## filtering and correct_mov are done in this function already
palate <- est_palate(corrected, coord[[1]], ref_idx, pl_idx, coord[[2]], coord[[3]])

```

# Feedback, Questions, and Contribution

This package is in active development.

If you have feature requests, bug reports, or questions, please either open an issue on Github or email Micah at [mehirsch\@bu.edu](mailto:mehirsch@bu.edu){.email}.

If you have ideas and would like to contribute to the project, please email Micah.

# License

MIT © 2025 SteppLab
